### 2016-08-23 syntax.tex
syntax

\heading{Preliminaries. } When~$\univ$ is a set and $n>0$ is an
integer, we let $[\univ]^n$ denote the set of all size-$n$ subsets
%\jnote{Multisets?} 
of~$\univ$, and $[[\univ]]^n$ for the set of all multisets
over~$\univ$ with total multiplicity~$n$. We write $x \getsr \univ$ to denote
sampling an element from~$\univ$ and assigning this to~$x$, and we
extend this notation to randomized algorithms.  When~$X$ is a string
over some alphabet~$\Sigma$, we write $|X|$ for the bitlength
of~$X$, relative to some fixed encoding.  When~$X$ is a multiset, we
overload the notation so that~$|X|$ is the total multiplicity of~$X$.

@@

\heading{Data structures.}  We being by defining a general primitive
that produces a representation of a given multiset, and provides a
mechanism to answer a specified collection of questions about that
multiset.  Informally, the representation should serve as a good
approximation of the multiset, at least with respect to the allowed
questions.
\begin{definition}[Data structure] \rm
Fix a non-empty set~$\univ$. A \emph{data structure} (over universe~$\univ$)
is a tuple $\setprim=(\Sigma,\Gamma,\calS,\calQ,\Rep, \Qry)$ where:
\begin{itemize}
\item $\Sigma$ is the \emph{representation alphabet}.
%\jnote{Is there any reason not to simply fix this to $\{0,1\}$?}
\item $\Gamma$ is the set of possible \emph{query answers}.
%\jnote{Do we need this for any reason?}
\item $\calS$ is a set  of multisets over universe~$\univ$.
\item $\mathcal{Q}=\{q \colon \mathcal{S} \to \Gamma\}$ is a set
    of supported \emph{queries}.
\item $\Rep \colon \mathcal{S} \to \Sigma^* \times \bits^*
    \times \bits^*$ is a randomized \emph{representation
    algorithm}, taking as input $S \in \mathcal{S}$, and
    outputting a \emph{representation} $M \in \Sigma^*$,
    \emph{private side information} $\privaux \in \bits^*$, and
    \emph{public side information} $\pubaux \in \bits^*$.  We write
    $(M,\pubaux,\privaux \getsr \Rep(S)$ as shorthand for this execution.
\item $\Qry\colon \Sigma^* \times \bits^* \times \bits^* \times \calQ \to
    \Gamma$ is a deterministic \emph{query-evaluation}
    algorithm, taking as input $M\in\Sigma^*$, $\pubaux,\privaux \in
    \bits^*$ and $q \in \calQ$, and outputting
    an \emph{answer} $a \in \Gamma$.  We write $a \gets
    \Qry(M,\pubaux,\privaux,q)$ for this execution.
\end{itemize}
\hfill\dqed
\end{definition}
%\tsnote{I implemented this change.}\tsnote{I'm now leaning towards
%having $\Rep$ output~$M$ and explicit pubic/private auxillary data.
%Pushing secrets into~$M$ makes it more difficult to capture (for
%example) the Neidermeier et al.\ attacks in our privacy notions; see
%Figure~\ref{fig:privacy}.  In those attacks, the adversary would get
%the bitarray (which one naturally thinks of as~$M$) but not the
%secret keys for the PRFs/hash functions.  You could say that the
%bitarray is put in $\aux$ and that~$M$ \emph{only} contains the
%secret keys, but this is unnatural.} 
All of the data structures that
we will consider have the property that~$|M|$ does not depend on the
coins of~$\Rep$. Thus we assume that~$|M|$ depends only on~$|S|$. In
general this need not be true, but the assumption does not restrict
our syntax from capturing natural constructions (e.g., those used in
practice), and it makes it easier to state our security notions.

@@

\heading{Set-multiplicity Data Structures. } Our particular focus
will be on data structures that support multiplicity queries, of
which set-membership queries are a special case.
\begin{definition} A \emph{set-multiplicity data structure} is a data
structure with $\Gamma=\mathbb{N}$ and
where $\calQ=\{q_x \,|\, x \in \univ\}$ where $q_x(S)$ is defined to
be the multiplicity of~$x$ in the multiset~$S$. \hfill\dqed
\end{definition}

@@

As a simple extension, consider $\Sigma=\mathbb{N}$ and modify the
basic Bloom filter so that it that stores counters, rather than
bits, at each position in an array.  In particular, for each~$x \in
S$ the $\Rep$~algorithm increments the counters stored at positions
$h_1(x), h_2(x), \ldots, h_k(x)$, returns the final array of
counters as~$M$.  The $\Qry$-algorithm remains the same. Such a
construction gives a static version of a counting Bloom
filter~\cite{xxx}.  We will capture dynamic versions, as well as
count-min sketches~\cite{xxx}, scalable Bloom filters~\cite{xxx},
etc., in a later section.

A Bloom filter with secret hash functions is captured by setting $\privaux=(h_1,h_2,\ldots,h_k)$, and setting $\pubaux=\emptystring$.  Related constructions \tsnote{Name them?} that use a secret key~$K$, but not for hashing, are captured setting $\privaux=K, \pubaux=(h_1, \ldots, h_k)$.

>>>>

\heading{Preliminaries. } When~$\univ$ is a set and $n>0$ is an
integer, we let $[\univ]^n$ denote the set of all size-$n$ subsets
of~$\univ$, and $[[\univ]]^n$ for the set of all multisets
over~$\univ$ with total multiplicity~$n$. \tsnote{We may not need all
  of that.} We write $x \getsr \univ$ to denote
sampling an element from~$\univ$ and assigning this to~$x$, and we
extend this notation to randomized algorithms.  When~$X$ is a string
over some alphabet~$\Sigma$, we write $|X|$ for the bitlength
of~$X$, relative to some fixed encoding.  When~$X$ is a multiset, we
overload the notation so that~$|X|$ is the total multiplicity of~$X$.

@@

\heading{Data structures.}  We being by defining a general primitive
that produces a representation of a given multiset, and provides a
mechanism to answer a specified collection of questions about that
multiset. 
%Informally, the representation should serve as a good
%approximation of the multiset, at least with respect to the allowed queries.
\begin{definition}[Data structure] \rm
Fix a non-empty set~$\univ$. A \emph{data structure} (over universe~$\univ$)
is a tuple $\setprim=(\Sigma,\Gamma,\calS,\calQ,\Rep, \Qry)$ where:
\begin{itemize}
\item $\Sigma$ is the \emph{representation alphabet}.
%\jnote{Is there any reason not to simply fix this to $\{0,1\}$?}
\item $\Gamma$ is the set of possible \emph{query answers}.
%\jnote{Do we need this for any reason?}
\item $\calS$ is a set  of multisets over universe~$\univ$.
\item $\mathcal{Q}=\{q \colon \mathcal{S} \to \Gamma\}$ is a set
    of supported \emph{queries}.
\item $\Rep \colon \mathcal{S} \to \Sigma^* \times \bits^*
    \times \bits^*$ is a randomized \emph{representation
    algorithm}, taking as input $S \in \mathcal{S}$, and
    outputting a \emph{representation} $M \in \Sigma^*$,
    \emph{private side information} $\privaux \in \bits^*$, and
    \emph{public side information} $\pubaux \in \bits^*$.  We write
    $(M,\pubaux,\privaux) \getsr \Rep(S)$ as shorthand for this execution.
\item $\Qry\colon \Sigma^* \times \bits^* \times \bits^* \times \calQ \to
    \Gamma$ is a deterministic \emph{query-evaluation}
    algorithm, taking as input $M\in\Sigma^*$, $\pubaux,\privaux \in
    \bits^*$ and $q \in \calQ$, and outputting
    an \emph{answer} $a \in \Gamma$.  We write $a \gets
    \Qry(M,\pubaux,\privaux,q)$ for this execution.
\end{itemize}
\hfill\dqed
\end{definition}
%\tsnote{I implemented this change.}\tsnote{I'm now leaning towards
%having $\Rep$ output~$M$ and explicit pubic/private auxillary data.
%Pushing secrets into~$M$ makes it more difficult to capture (for
%example) the Neidermeier et al.\ attacks in our privacy notions; see
%Figure~\ref{fig:privacy}.  In those attacks, the adversary would get
%the bitarray (which one naturally thinks of as~$M$) but not the
%secret keys for the PRFs/hash functions.  You could say that the
%bitarray is put in $\aux$ and that~$M$ \emph{only} contains the
%secret keys, but this is unnatural.} 
All of the data structures that
we will consider have the property that~$|M|$ does not depend on the
coins of~$\Rep$. Thus we assume that~$|M|$ depends only on~$|S|$. In
general this need not be true, but the assumption does not restrict
our syntax from capturing natural constructions (e.g., those used in
practice), and it makes it easier to state our security notions.

@@

\heading{Data Structures for Set-Multiplicity Queries. } Our particular focus
will be on data structures that support multiplicity queries, of
which set-membership queries are a special case.
\begin{definition} A \emph{set-multiplicity data structure} is a data
structure with $\Gamma=\mathbb{N}$ and
where $\calQ=\{q_x \,|\, x \in \univ\}$ where $q_x(S)$ is defined to
be the multiplicity of~$x$ in the multiset~$S$. \hfill\dqed
\end{definition}

@@

As a simple extension, consider $\Sigma=\mathbb{N}$ and modify the
basic Bloom filter so that it that stores counters, rather than
bits, at each position in an array.  In particular, for each~$x \in
S$ the $\Rep$~algorithm increments the counters stored at positions
$h_1(x), h_2(x), \ldots, h_k(x)$, returns the final array of
counters as~$M$.  The $\Qry$-algorithm remains the same. Such a
construction gives a static version of a counting Bloom
filter~\cite{xxx}.  We will capture dynamic versions, as well as
count-min sketches~\cite{xxx}, scalable Bloom filters~\cite{xxx},
etc., in a later section.\tsnote{If we ultimately do!}

A Bloom filter with secret hash functions is captured by setting
$\privaux=(h_1,h_2,\ldots,h_k)$, and setting $\pubaux=\emptystring$.
Related constructions that use a secret key~$K$, but not for hashing (e.g. the PRP-based
construction from NY), are captured setting $\privaux=K, \pubaux=(h_1, \ldots, h_k)$.
